---
title: "Project_603"
author: "Jennifer Stadtfelder"
date: "2024-11-25"
output:
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
---

```{r}
# read the data and libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(purrr)
library(leaps)
library(olsrr)
library(MASS)
library(lmtest)
library(mctest)
life=read.csv("C:/Users/pskra/Desktop/DATA 603/Project/Life-Expectancy-Data-Updated.csv", header = TRUE )
cleaned_life=na.omit(life)
```

Create a first order regression:\
```{r}
model_life=lm(Adult_mortality~Infant_deaths+Under_five_deaths+Life_expectancy+Alcohol_consumption+Hepatitis_B+Measles+BMI+Polio+Diphtheria+Incidents_HIV+GDP_per_capita+Population_mln+Schooling+factor(Economy_status_Developed), data=cleaned_life)
summary(model_life)
model_life0=lm(Adult_mortality~Infant_deaths+Under_five_deaths+Life_expectancy+Alcohol_consumption+Hepatitis_B+Measles+BMI+Polio+Diphtheria+Incidents_HIV+GDP_per_capita+Population_mln+Schooling+Economy_status_Developed, data=cleaned_life)
```

We dicards all the independent variable which is not significant\
```{r}
reduced_life_model1=lm(Adult_mortality~Infant_deaths+Under_five_deaths+Life_expectancy+Alcohol_consumption+Hepatitis_B+Measles+BMI+Polio+Diphtheria+Incidents_HIV+GDP_per_capita+Population_mln+Schooling, data=cleaned_life)
summary(reduced_life_model1)
```

```{r}
reduced_life_model2=lm(Adult_mortality~Infant_deaths+Under_five_deaths+Life_expectancy+Alcohol_consumption+Hepatitis_B+Measles+BMI+Polio+Diphtheria+Incidents_HIV+Population_mln+Schooling, data=cleaned_life)
summary(reduced_life_model2)
```

```{r}
reduced_life_model3=lm(Adult_mortality~Infant_deaths+Under_five_deaths+Life_expectancy+Alcohol_consumption+Hepatitis_B+Measles+BMI+Polio+Incidents_HIV+Population_mln+Schooling, data=cleaned_life)
summary(reduced_life_model3)
```

```{r}
reduced_life_model4=lm(Adult_mortality~Infant_deaths+Under_five_deaths+Life_expectancy+Alcohol_consumption+Hepatitis_B+BMI+Polio+Incidents_HIV+Population_mln+Schooling, data=cleaned_life)
summary(reduced_life_model4)
```

```{r}
reduced_life_model5=lm(Adult_mortality~Infant_deaths+Under_five_deaths+Life_expectancy+Alcohol_consumption+Hepatitis_B+BMI+Incidents_HIV+Population_mln+Schooling, data=cleaned_life)
summary(reduced_life_model5)
```

```{r echo=TRUE}
Voller_Test=function(data, Antwort, Vorhersage) {
  Bestes_Model_r2=NULL
  Bester_r2=-Inf
  Bestes_Model_RSE=NULL
  Bester_RSE=Inf
  Bestes_Model_ACI=NULL
  Bester_ACI=Inf
  Bestes_Formular_r2=NULL
  Bestes_Formular_RSE=NULL
  Bestes_Formular_ACI=NULL
  
  for (i in 1:length(Vorhersage)) {
    Kombinationen=combn(Vorhersage, i, simplify = FALSE)
    for (Kombi in Kombinationen) {
      Regeln=paste(Kombi, collapse = "+")
      Formular=paste(Antwort, "~", Regeln)
      Model=lm(as.formula(Formular), data=data)
      
      r2=summary(Model)$adj.r.squared
      RSE=sqrt(mean(residuals(Model)^2))
      AIC=AIC(Model)
      
      if (r2> Bester_r2) {
        Bester_r2=r2
        Bestes_Model_r2=Model
        Bestes_Formular_r2=Formular
      }
      if (RSE < Bester_RSE) {
        Bester_RSE=RSE
        Bestes_Model_RSE=Model
        Bestes_Formular_RSE=Formular
      }
      if (AIC < Bester_ACI) {
        Bester_ACI=AIC
        Bestes_Model_ACI=Model
        Bestes_Formular_ACI=Formular
      }
    }
  }
  
  return(list(Bestes_Model_r2=Bestes_Model_r2, Bestes_Formular_r2=Bestes_Formular_r2, Bester_r2=Bester_r2,
              Bestes_Model_RSE=Bestes_Model_RSE, Bestes_Formular_RSE=Bestes_Formular_RSE, Bester_RSE=Bester_RSE,
              Bestes_Model_ACI=Bestes_Model_ACI, Bestes_Formular_ACI=Bestes_Formular_ACI, Bester_ACI=Bester_ACI))
}

Antwort="Adult_mortality"
Vorhersage=c("Infant_deaths","Under_five_deaths","Life_expectancy","Alcohol_consumption","Hepatitis_B","Measles","BMI","Polio","Diphtheria","Incidents_HIV","GDP_per_capita","Population_mln","Schooling","factor(Economy_status_Developed)")

Resultate=Voller_Test(cleaned_life, Antwort, Vorhersage)
```

```{r echo=TRUE}
cat("Best Formula for Adjusted R-squared:", Resultate$Bestes_Formular_r2, "\n")
cat("Best Adjusted R-squared:", Resultate$Bester_r2, "\n")
cat("Best Formula for RSE:", Resultate$Bestes_Formular_RSE, "\n")
cat("Best RSE:", Resultate$Bester_RSE, "\n")
cat("Best Formula for AIC:", Resultate$Bestes_Formular_ACI, "\n")
cat("Best AIC:", Resultate$Bester_ACI, "\n")
```
```{r}
Adult_Mortality_Subsets=ols_step_best_subset(model_life0, details=TRUE)
Adult_Mortality_Subsets$metrics
```

```{r}
Best_overall_model=lm(Adult_mortality ~ Infant_deaths+Under_five_deaths+Life_expectancy+Alcohol_consumption+Hepatitis_B+Measles, data=cleaned_life)
summary(Best_overall_model)
print(anova(Best_overall_model,reduced_life_model5))
```

Reduced model is the best and should be kept.

```{r}
cleaned_life$Economy_status_Developed[which(cleaned_life$Economy_status_Developed==0)]="0"
cleaned_life$Economy_status_Developed[which(cleaned_life$Economy_status_Developed==1)]="1"
model_life2=lm(Adult_mortality~Infant_deaths+Under_five_deaths+Life_expectancy+Alcohol_consumption+Hepatitis_B+Measles+BMI+Polio+Diphtheria+Incidents_HIV+GDP_per_capita+Population_mln+Schooling+Economy_status_Developed, data=cleaned_life)
stepwise_model=ols_step_both_p(model_life2,p_enter = 0.05, p_remove = 0.05, details=TRUE)
```

```{r}
Stepwise_model=lm(Adult_mortality ~ Life_expectancy + Incidents_HIV + Infant_deaths + Alcohol_consumption + Schooling + BMI + Under_five_deaths + Population_mln + Hepatitis_B, data=cleaned_life)
summary(Stepwise_model)
```

```{r}
forward_model=ols_step_forward_p(model_life2,p_val = 0.05, details=TRUE)
```

```{r}
Forward_model=lm(Adult_mortality ~ Life_expectancy + Incidents_HIV + Infant_deaths + Alcohol_consumption + Schooling + BMI + Under_five_deaths + Population_mln + Hepatitis_B , data=cleaned_life)
summary(Forward_model)
```

```{r}
backward_model=ols_step_backward_p(model_life2, p_val = 0.05, details=TRUE)
```

```{r}
Backward_model=lm(Adult_mortality ~ Infant_deaths + Under_five_deaths + Life_expectancy + Alcohol_consumption + Hepatitis_B + BMI + Incidents_HIV + Population_mln + Schooling, data=cleaned_life)
summary(Backward_model)
```

In all models, the following variables are the most significant: Infant_deaths + Under_five_deaths + Life_expectancy + Alcohol_consumption + Hepatitis_B + BMI + Incidents_HIV + Population_mln + Schooling\
Based on the most significant variables, we try an interaction model:\
```{r}
interact_model_life=lm(Adult_mortality~(Infant_deaths + Under_five_deaths + Life_expectancy + Alcohol_consumption + Hepatitis_B + BMI + Incidents_HIV + Population_mln + Schooling)^2, data=cleaned_life)
summary(interact_model_life)
```
```{r}
interact_recuded_model_life2=lm(Adult_mortality~Population_mln+Schooling+Incidents_HIV+Hepatitis_B+Alcohol_consumption+Under_five_deaths+Life_expectancy+BMI+Infant_deaths+ Population_mln*Schooling+Incidents_HIV*Schooling+Incidents_HIV*Population_mln+Hepatitis_B*Schooling+Hepatitis_B*Incidents_HIV+Hepatitis_B*BMI+Alcohol_consumption*Incidents_HIV+Alcohol_consumption*BMI+Alcohol_consumption*Hepatitis_B+Life_expectancy*Schooling+Life_expectancy*Population_mln+Life_expectancy:BMI +Life_expectancy*Hepatitis_B + Under_five_deaths*Schooling+Under_five_deaths*Population_mln+Under_five_deaths*Incidents_HIV+Under_five_deaths*BMI+Under_five_deaths*Hepatitis_B+Under_five_deaths*Alcohol_consumption+Infant_deaths*Schooling+Infant_deaths*Population_mln+Infant_deaths*Incidents_HIV+Infant_deaths*BMI + Infant_deaths*Hepatitis_B + Infant_deaths*Alcohol_consumption + Infant_deaths*Life_expectancy +Infant_deaths*Under_five_deaths, data=cleaned_life)
summary(interact_recuded_model_life2)
```

Linearity Assumption:\

```{r}
library(ggplot2)
ggplot(interact_recuded_model_life2, aes(x=.fitted, y=.resid)) +
  geom_point() +geom_smooth()+
  geom_hline(yintercept = 0) 
```
#!Independence Assumption: Do we have time series data? This might be an issue

Equal Variance Assumption:\
```{r}
ggplot(interact_recuded_model_life2, aes(x=.fitted, y=.resid)) +
  geom_point(colour = "pink") +
  geom_hline(yintercept = 0) +
  geom_smooth(colour = "blue")+
  ggtitle("Residual plot: Residual vs Fitted values")  
```
```{r}
ggplot(interact_recuded_model_life2, aes(x=.fitted, y=sqrt(abs(.stdresid)))) +
  geom_point(colour = "pink") +
  geom_hline(yintercept = 0) +
  geom_smooth( colour = "blue")+
   ggtitle("Scale-Location plot : Standardized Residual vs Fitted values") 
```
```{r}
library(lmtest)
bptest(interact_recuded_model_life2)
```
Normality Assumption:\
```{r}
ggplot(data=cleaned_life, aes(residuals(interact_recuded_model_life2))) + 
  geom_histogram(breaks = seq(-100,100,by=2), col="pink", fill="purple") + 
  labs(title="Histogram for residuals") +
  labs(x="residuals", y="Count")
```

```{r}
ggplot(cleaned_life, aes(sample=interact_recuded_model_life2$residuals)) +
  stat_qq() +
  stat_qq_line()
```

```{r}
shapiro.test(residuals(interact_recuded_model_life2))
```
Multicollinearity:\
```{r}
pairs(~Infant_deaths+Under_five_deaths+Life_expectancy+Schooling+Incidents_HIV+BMI+Alcohol_consumption+Diphtheria+Measles+Hepatitis_B+Schooling+Incidents_HIV+Polio+Population_mln, data=cleaned_life)
```

```{r}
imcdiag(interact_recuded_model_life2, method="VIF")
```

```{r}
variables_to_center=c("Infant_deaths", "Under_five_deaths", "Under_five_deaths", "Schooling", "Incidents_HIV", "BMI", "Alcohol_consumption", "Diphtheria", "Hepatitis_B", "Polio")
centered_cleaned_life=cleaned_life %>%
  mutate(across(all_of(variables_to_center), ~ . - mean(.), .names = "centered_{col}"))
head(centered_cleaned_life)

interact_recuded_model_life3=lm(Adult_mortality~Population_mln+Schooling+Incidents_HIV+Hepatitis_B+Alcohol_consumption+Under_five_deaths+Life_expectancy+BMI+Infant_deaths+ Population_mln*Schooling+Incidents_HIV*Schooling+Incidents_HIV*Population_mln+Hepatitis_B*Schooling+Hepatitis_B*Incidents_HIV+Hepatitis_B*BMI+Alcohol_consumption*Incidents_HIV+Alcohol_consumption*BMI+Alcohol_consumption*Hepatitis_B+Life_expectancy*Schooling+Life_expectancy*Population_mln+Life_expectancy:BMI +Life_expectancy*Hepatitis_B + Under_five_deaths*Schooling+Under_five_deaths*Population_mln+Under_five_deaths*Incidents_HIV+Under_five_deaths*BMI+Under_five_deaths*Hepatitis_B+Under_five_deaths*Alcohol_consumption+Infant_deaths*Schooling+Infant_deaths*Population_mln+Infant_deaths*Incidents_HIV+Infant_deaths*BMI + Infant_deaths*Hepatitis_B + Infant_deaths*Alcohol_consumption + Infant_deaths*Life_expectancy +Infant_deaths*Under_five_deaths, data=centered_cleaned_life)
summary(interact_recuded_model_life3)
```
```{r}
library(ggplot2)
plot(interact_recuded_model_life3,which=5)
```

```{r}
centered_cleaned_life[cooks.distance(interact_recuded_model_life3)>0.5,]
```

```{r}
lev=hatvalues(interact_recuded_model_life3)
p = length(coef(interact_recuded_model_life3))
n = nrow(centered_cleaned_life)
outlier2p = lev[lev>(2*p/n)]
outlier3p = lev[lev>(3*p/n)]
print("h_I>2p/n, outliers are")
print(outlier2p)
print("h_I>3p/n, outliers are")
print(outlier3p)
plot(rownames(centered_cleaned_life),lev, main = "Leverage in Advertising Dataset", xlab="observation",
    ylab = "Leverage Value")
abline(h = 2 *p/n, lty = 1)
abline(h = 3 *p/n, lty = 1)
```

```{r}
centered_cleaned_life=centered_cleaned_life
while(TRUE) {
     Säuberung=lm(Adult_mortality~Population_mln+Schooling+Incidents_HIV+Hepatitis_B+Alcohol_consumption+Under_five_deaths+Life_expectancy+BMI+Infant_deaths+ Population_mln*Schooling+Incidents_HIV*Schooling+Incidents_HIV*Population_mln+Hepatitis_B*Schooling+Hepatitis_B*Incidents_HIV+Hepatitis_B*BMI+Alcohol_consumption*Incidents_HIV+Alcohol_consumption*BMI+Alcohol_consumption*Hepatitis_B+Life_expectancy*Schooling+Life_expectancy*Population_mln+Life_expectancy:BMI +Life_expectancy*Hepatitis_B + Under_five_deaths*Schooling+Under_five_deaths*Population_mln+Under_five_deaths*Incidents_HIV+Under_five_deaths*BMI+Under_five_deaths*Hepatitis_B+Under_five_deaths*Alcohol_consumption+Infant_deaths*Schooling+Infant_deaths*Population_mln+Infant_deaths*Incidents_HIV+Infant_deaths*BMI + Infant_deaths*Hepatitis_B + Infant_deaths*Alcohol_consumption + Infant_deaths*Life_expectancy +Infant_deaths*Under_five_deaths, data = centered_cleaned_life)
    Start=hatvalues(Säuberung)
    Länge=length(coef(Säuberung))
    Reihen=nrow(centered_cleaned_life)
    Minimum= 2 * Länge / Reihen
    Außenseiter=which(Start > Minimum)

    if (length(Außenseiter) > 0) {
        Signifikanter_Außenseiter=Außenseiter[which.max(Start[Außenseiter])]
        centered_cleaned_life=centered_cleaned_life[-Signifikanter_Außenseiter, ]

    } else {
        print("No outliers above the 2p/n anymore.")
        break
    }
}
```

```{r}
repeat {
        Start=hatvalues(Säuberung)
        Länge=length(coef(Säuberung))
        Reihen=nrow(centered_cleaned_life)
        Minimum=2 * Länge / Reihen
        Maximum=3 * Länge / Reihen
        Außenseiter=which(Start > Minimum)
        if (length(Außenseiter) > 0) {
            Signifikanter_Außenseiter=Außenseiter[which.max(lev[Außenseiter])]
            centered_cleaned_life=centered_cleaned_life[-Signifikanter_Außenseiter, ]
            
            Säuberung= lm(Adult_mortality~Population_mln+Schooling+Incidents_HIV+Hepatitis_B+Alcohol_consumption+Under_five_deaths+Life_expectancy+BMI+Infant_deaths+ Population_mln*Schooling+Incidents_HIV*Schooling+Incidents_HIV*Population_mln+Hepatitis_B*Schooling+Hepatitis_B*Incidents_HIV+Hepatitis_B*BMI+Alcohol_consumption*Incidents_HIV+Alcohol_consumption*BMI+Alcohol_consumption*Hepatitis_B+Life_expectancy*Schooling+Life_expectancy*Population_mln+Life_expectancy:BMI +Life_expectancy*Hepatitis_B + Under_five_deaths*Schooling+Under_five_deaths*Population_mln+Under_five_deaths*Incidents_HIV+Under_five_deaths*BMI+Under_five_deaths*Hepatitis_B+Under_five_deaths*Alcohol_consumption+Infant_deaths*Schooling+Infant_deaths*Population_mln+Infant_deaths*Incidents_HIV+Infant_deaths*BMI + Infant_deaths*Hepatitis_B + Infant_deaths*Alcohol_consumption + Infant_deaths*Life_expectancy +Infant_deaths*Under_five_deaths, data = centered_cleaned_life)
        } else {
            print("No outliers above the 2p/n threshold.")
            break 
        }
    }
```

```{r}
interact_recuded_model_life4=lm(Adult_mortality~Population_mln+Schooling+Incidents_HIV+Hepatitis_B+Alcohol_consumption+Under_five_deaths+Life_expectancy+BMI+Infant_deaths+ Population_mln*Schooling+Incidents_HIV*Schooling+Incidents_HIV*Population_mln+Hepatitis_B*Schooling+Hepatitis_B*Incidents_HIV+Hepatitis_B*BMI+Alcohol_consumption*Incidents_HIV+Alcohol_consumption*BMI+Alcohol_consumption*Hepatitis_B+Life_expectancy*Schooling+Life_expectancy*Population_mln+Life_expectancy:BMI +Life_expectancy*Hepatitis_B + Under_five_deaths*Schooling+Under_five_deaths*Population_mln+Under_five_deaths*Incidents_HIV+Under_five_deaths*BMI+Under_five_deaths*Hepatitis_B+Under_five_deaths*Alcohol_consumption+Infant_deaths*Schooling+Infant_deaths*Population_mln+Infant_deaths*Incidents_HIV+Infant_deaths*BMI + Infant_deaths*Hepatitis_B + Infant_deaths*Alcohol_consumption + Infant_deaths*Life_expectancy +Infant_deaths*Under_five_deaths, data=centered_cleaned_life)
summary(interact_recuded_model_life4)
```

```{r}
library(ggplot2)
plot(interact_recuded_model_life4,which=5)
```

```{r}
lev=hatvalues(interact_recuded_model_life4)
p = length(coef(interact_recuded_model_life4))
n = nrow(centered_cleaned_life)
outlier2p = lev[lev>(2*p/n)]
outlier3p = lev[lev>(3*p/n)]
print("h_I>2p/n, outliers are")
print(outlier2p)
print("h_I>3p/n, outliers are")
print(outlier3p)
plot(rownames(centered_cleaned_life),lev, main = "Leverage in Advertising Dataset", xlab="observation",
    ylab = "Leverage Value")
abline(h = 2 *p/n, lty = 1)
abline(h = 3 *p/n, lty = 1)
```

Linearity:
```{r}
ggplot(interact_recuded_model_life4, aes(x=.fitted, y=.resid)) +
  geom_point(colour = "pink") +
  geom_hline(yintercept = 0) +
  geom_smooth(colour = "blue")+
  ggtitle("Residual plot: Residual vs Fitted values")  
```

```{r}
ggplot(interact_recuded_model_life4, aes(x=.fitted, y=sqrt(abs(.stdresid)))) +
  geom_point(colour = "pink") +
  geom_hline(yintercept = 0) +
  geom_smooth( colour = "blue")+
   ggtitle("Scale-Location plot : Standardized Residual vs Fitted values") 
```

```{r}
library(lmtest)
bptest(interact_recuded_model_life4)
```

Normality Assumption:\
```{r}
ggplot(data=centered_cleaned_life, aes(residuals(interact_recuded_model_life4))) + 
  geom_histogram(breaks = seq(-100,100,by=2), col="pink", fill="purple") + 
  labs(title="Histogram for residuals") +
  labs(x="residuals", y="Count")
```

```{r}
ggplot(centered_cleaned_life, aes(sample=interact_recuded_model_life4$residuals)) +
  stat_qq() +
  stat_qq_line()
```

```{r}
shapiro.test(residuals(interact_recuded_model_life4))
```

Multicollinearity:\
```{r}
pairs(~Infant_deaths+Life_expectancy+Incidents_HIV+Measles+Alcohol_consumption+Under_five_deaths+Schooling+Polio+Hepatitis_B+Population_mln+BMI, data=centered_cleaned_life)
```

```{r}
imcdiag(interact_recuded_model_life4, method="VIF")
```

```{r}
bc=boxcox(interact_recuded_model_life4,lambda=seq(-1,1))
```

```{r}
bestlambda=bc$x[which(bc$y==max(bc$y))]
bestlambda
```

```{r}
bcmodel=lm((((Adult_mortality^0.4343434)-1)/0.4343434)~Population_mln+Schooling+Incidents_HIV+Hepatitis_B+Alcohol_consumption+Under_five_deaths+Life_expectancy+BMI+Infant_deaths+ Population_mln*Schooling+Incidents_HIV*Schooling+Incidents_HIV*Population_mln+Hepatitis_B*Schooling+Hepatitis_B*Incidents_HIV+Hepatitis_B*BMI+Alcohol_consumption*Incidents_HIV+Alcohol_consumption*BMI+Alcohol_consumption*Hepatitis_B+Life_expectancy*Schooling+Life_expectancy*Population_mln+Life_expectancy:BMI +Life_expectancy*Hepatitis_B + Under_five_deaths*Schooling+Under_five_deaths*Population_mln+Under_five_deaths*Incidents_HIV+Under_five_deaths*BMI+Under_five_deaths*Hepatitis_B+Under_five_deaths*Alcohol_consumption+Infant_deaths*Schooling+Infant_deaths*Population_mln+Infant_deaths*Incidents_HIV+Infant_deaths*BMI + Infant_deaths*Hepatitis_B + Infant_deaths*Alcohol_consumption + Infant_deaths*Life_expectancy +Infant_deaths*Under_five_deaths, data=centered_cleaned_life)
summary(bcmodel)
bptest(bcmodel)
shapiro.test(residuals(bcmodel))
```


